---
title: "Multi Factor Analysis"
author: "MRC CSC Bioinformatics Core"
date: "08 August 2016"
output: html_document
Website: http://bioinformatics.csc.mrc.ac.uk/training.html
---

1.  Using likelihood ratio test instead of the wald test,

a) Find the number of genes that are changed because of condition at FDR 0.05 irrespective of fold change.

b) Find the number of genes that are changed because of batch at FDR 0.05 irrespective of fold change.

```{r, echo=T,warnings=F,collapse=T,eval=FALSE}

ddsLRT<-DESeqDataSetFromMatrix(countData=AllCounts,colData=cData,design=~batch+condition)

# LRT analysis for the condition effect
ddsLRT_con <- DESeq(ddsLRT, test="LRT", full=~batch+condition, reduced=~batch) 

resddsLRT_con<-results(ddsLRT_con)

summary(resddsLRT_con,alpha=0.05)

resLRTorder<-resddsLRT_con[order(resddsLRT_con$padj),]

```


2. Perform multi factor analysis by adding batch as factor to design.

a) Find out number of genes that are changed because of batch as opposed to because of Group.

```{r, echo=T,warnings=F,collapse=T,eval=FALSE}

# LRT analysis for the batch effect


ddsLRT_batch <- DESeq(ddsLRT, test="LRT", full=~batch+condition, reduced=~condition)

resddsLRT_batch<-results(ddsLRT_batch)

summary(resddsLRT_batch,alpha=0.05)

```


