---
title: "Multi Factor Analysis"
author: "MRC CSC Bioinformatics Core"
date: "08 August 2016"
output: html_document
Website: http://bioinformatics.csc.mrc.ac.uk/training.html
---

1. Use contrasts to specify the comparison you wish to make. Run comparisons between KOa vs FFa and KOb vs FFa. 

2. Perform multi factor analysis by adding batch as factor to design.

3. Repeat above analysis using likelihood ratio test instead of the wald test,

a) Find the number of genes that are changed because of condition at FDR 0.05.


```{r, echo=T,warnings=F,collapse=T,eval=FALSE}

# LRT analysis for the batch effect


ddsLRT_batch <- DESeq(ddsLRT, test="LRT", full=~batch+condition, reduced=~condition)

resddsLRT_batch<-results(ddsLRT_batch)

summary(resddsLRT_batch,alpha=0.05)

```

b) Find the number of genes that are changed because of batch at FDR 0.05.

```{r, echo=T,warnings=F,collapse=T,eval=FALSE}

# LRT analysis for the batch effect


ddsLRT_batch <- DESeq(ddsLRT, test="LRT", full=~batch+condition, reduced=~condition)

resddsLRT_batch<-results(ddsLRT_batch)

summary(resddsLRT_batch,alpha=0.05)

```



3.  Using likelihood ratio test instead of the wald test,

a) Find the number of genes that are changed because of condition at FDR 0.05 irrespective of fold change.

b) Find the number of genes that are changed because of batch at FDR 0.05 irrespective of fold change.

```{r, echo=T,warnings=F,collapse=T,eval=FALSE}

ddsLRT<-DESeqDataSetFromMatrix(countData=AllCounts,colData=cData,design=~batch+condition)

# LRT analysis for the condition effect
ddsLRT_con <- DESeq(ddsLRT, test="LRT", full=~batch+condition, reduced=~batch) 

resddsLRT_con<-results(ddsLRT_con)

summary(resddsLRT_con,alpha=0.05)

resLRTorder<-resddsLRT_con[order(resddsLRT_con$padj),]

```





